### 2019-08-13 chapter_02_open_science.tex
Change figure captions and positions
More coreections and improved figures in exp chapter

It is normal that Mathworks, the company which makes money with Matlab, spreads
\fnref{https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
against Python and its scientific ecosystem}.  Some of them are valid, for
example the issue of the lively, huge and thus complicated scientific ecosystem
that we have already discussed. Other arguments are unfair or no longer valid
nowadays\footnote{See also \url{http://www.pyzo.org/python\_vs\_matlab.html} or
\url{https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen}}.
%
The Python scientific ecosystem can now compete in nearly all domains, but this
was not the case typically five years ago.
%
It would be interesting to estimate what open-source solutions could benefit
from using the amount paid by public research institutes. How many
developers could be paid to improve the alternative open-source solutions?
%
This question raises the issue of funding scientific open-source software.

@@

% TODO: perhaps include the commented out sections in the beginning of the chapter
% pa: I am really uncomfortable with this presentation which gives the
% impression that we think that the arguments against sharing are always wrong.
%
% We have to make it clear that it is not the case and that we do respect how
% other people think and work.
% av: I have commented out specific counter arguments, instead added two
% sentences after the bullets
\begin{itemize}
\item \textit{``We do not share to keep a comparative advantage.''}

@@

Finally, several scientists in the fluid dynamics community have used
open-source methods in their work for a long time.
%
There are a lot of open-source research codes written with proprietary tools
(typically written in Matlab, see for example
\fnref{http://pivlab.blogspot.de/}{PIVlab},
\fnref{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVmat} and
\fnref{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
%
Let us first say that these codes are a great wealth for the community, as they
can be used as they are, and other open-source research projects can reuse 
the ideas expressed in these codes.
%
Although the codes are meant to be free, users need to pay to obtain the
compilers/interpreters which execute them. 
%
% Do they realize that when they work on these programs,
Ironically, when people develop programs using these tools, unintentionally
they work (often, for free) for the company that sells the proprietary tool.

@@

\section{FluidDyn project}
%: Breaking the status quo}
% \section{Implementation and architecture}
% TODO: a lot of self-plagiarism below - needs rewrite
We have seen that there is a strong dynamics in play around the use of
computers (in particular with the web) and that this creates very efficient
tools and methods for collective work and software development.
%
The potential of these tools is not fully exploited yet in the field of fluid
mechanics.

>>>>

It is normal that Mathworks, the company which makes money with Matlab, spreads
\fnref{https://web.archive.org/web/20190807004347/https://www.mathworks.com/products/matlab/matlab-vs-python.html}{arguments
against Python and its scientific ecosystem}.  Some of them are valid, for
example the issue of the lively, huge and thus complicated scientific ecosystem
that we have already discussed. Other arguments are unfair or no longer valid
nowadays\footnote{See also \url{http://www.pyzo.org/python\_vs\_matlab.html} or
\url{https://www.linkedin.com/pulse/matlab-vs-python-jan-rhebergen}}.
%
The Python scientific ecosystem can now compete in nearly all domains, but this
was not the case typically five years ago.
%
It would be interesting to estimate what open-source solutions could benefit
from using the amount paid by public research institutes. How many
developers could be paid to improve the alternative open-source solutions?
%
This question raises the issue of funding scientific open-source software.

@@

% pa: I am really uncomfortable with this presentation which gives the
% impression that we think that the arguments against sharing are always wrong.
%
% We have to make it clear that it is not the case and that we do respect how
% other people think and work.
% av: I have commented out specific counter arguments, instead added two
% sentences after the bullets
\begin{itemize}
\item \textit{``We do not share to keep a comparative advantage.''}

@@

Finally, several scientists in the fluid dynamics community have used
open-source methods in their work for a long time.
%
There are a lot of open-source research codes written with proprietary tools
(typically written in Matlab, see for example
\fnref{http://pivlab.blogspot.de/}{PIVlab},
\fnref{http://servforge.legi.grenoble-inp.fr/projects/soft-uvmat/}{UVMAT} and
\fnref{https://de.mathworks.com/matlabcentral/fileexchange/10902-pivmat-4-00}{PIVmat}).
%
Let us first say that these codes are a great wealth for the community, as they
can be used as they are, and other open-source research projects can reuse
the ideas expressed in these codes.
%
Although the codes are meant to be free, users need to pay to obtain the
compilers/interpreters which execute them.
%
% Do they realize that when they work on these programs,
Ironically, when people develop programs using these tools, unintentionally
they work (often, for free) for the company that sells the proprietary tool.

@@

\section{FluidDyn project}
%: Breaking the status quo}
% \section{Implementation and architecture}
We have seen that there is a strong dynamics in play around the use of
computers (in particular with the web) and that this creates very efficient
tools and methods for collective work and software development.
%
The potential of these tools is not fully exploited yet in the field of fluid
mechanics.
